Linear Regret Bounds for Polyhedral Surrogates

### 2021-05-19 notes/regret-bound.tex
read regret upper bound proof.

@@

\begin{theorem}[Hoffman constant]
  \label{thm:hoffman}
  Given a matrix $A\in\reals^{m\times n}$, there exists some $H(A)\geq 0$, called the \emph{Hoffman constant} (with respect to $\|\cdot\|_\infty$), such that for all $b\in\reals^m$ and all $x\in\reals^n$,
  \begin{equation}
    \label{eq:hoffman}
    d_\infty(x,S(A,b)) \leq H(A) \|(A x - b)_+\|_\infty~,
  \end{equation}
  where $S(A,b) = \{x\in\reals^n \mid A x \leq b\}$ and $(u)_+ \defeq \max(u,0)$ component-wise.
\end{theorem}

@@

\begin{lemma}[Regret bound for fixed $p$]
  \label{lem:regret-bound-fixed-p}
  Let surrogate $L:\reals^d \to \reals^\Y_+$ and discrete loss $\ell:\R\to\reals^\Y_+$ be given.
  Let $\psi:\reals^d\to\R$ be an $\epsilon$-separated link with respect to $\Gamma \defeq \prop{L}$ and $\gamma \defeq \prop{\ell}$ for some $\epsilon>0$.
  For fixed $p\in\simplex$, there exists some $\alpha\geq 0$ such that
  \begin{equation}
    \label{eq:surrogate-regret-bound-fixed}
    \regret{\ell}{\psi(u)}{p}
    \leq
    \alpha \cdot \regret{L}{u}{p}~,
    % \inprod{p}{\ell(\psi(u))} - \risk{\ell}(p)
    % \leq
    % \alpha \cdot \left( \inprod{p}{L(u)} - \risk{L}(p) \right)~,
  \end{equation}
  for all $u\in\reals^d$.
\end{lemma}
\begin{proof}
  Since $L$ is polyhedral, there exist $a_1,\ldots,a_m \in \reals^d$ and $c\in\reals^m$ such that we may write $\inprod{p}{L(u)} = \max_{1\leq j\leq m} a_j \cdot u + c_j$.
  Let $A \in \reals^{m\times d}$ be the matrix with rows $a_j$, and let $b = \risk{L}(p)\ones - c$, where $\ones\in\reals^m$ is the all-ones vector.
  Then we have
  \begin{align*}
    S(A,b)
    &\defeq \{u\in\reals^d \mid A u \leq b\}
    \\
    &= \{u\in\reals^d \mid A u + c \leq \risk{L}(p)\ones\}
    \\
    &= \{u\in\reals^d \mid \forall i\, (A u + c)_i \leq \risk{L}(p)\}
    \\
    &= \{u\in\reals^d \mid \max_i \;(A u + c)_i \leq \risk{L}(p)\}
    \\
    &= \{u\in\reals^d \mid \inprod{p}{L(u)} \leq \risk{L}(p)\}
    \\
    & = \Gamma(p)~.
  \end{align*}
  Similarly, we have $\max_i\; (A u - b)_i = \inprod{p}{L(u)} - \risk{L}(p) = \regret{L}{u}{p} \geq 0$.
  Thus,
  \begin{align*}
    \|(Au - b)_+\|_\infty
    &= \max_i\; ((Au - b)_+)_i
    \\
    &= \max((Au - b)_1,\ldots,(Au - b)_m, 0)
    \\
    &= \max(\max_i\; (Au - b)_i, \, 0)
    \\
    &= \max_i\; (Au - b)_i
    % \\
    % &= \inprod{p}{L(u)} - \risk{L}(p)
    \\
    &= \regret{L}{u}{p}~.
  \end{align*}

  Now applying Theorem~\ref{thm:hoffman}, we have
  \begin{align*}
    d_\infty(u,\Gamma(p))
    &\leq H(A) \|(Au-b)_+\|_\infty
    \\
    &= H(A) \regret{L}{u}{p}~.
  \end{align*}
  Let $\beta \defeq \max_{r\in\R} \regret{\ell}{r}{p}\geq 0$, which is well-defined as $\R$ is a finite set.
  To complete the proof, define $\alpha = \beta H(A)/\epsilon$.
  Given $u\in\reals^d$, if $\regret{\ell}{\psi(u)}{p} = 0$, we are done.
  Otherwise, we have from $\epsilon$-separation that $d_\infty(u,\Gamma(p)) > \epsilon$.
  The above then gives $H(A) \regret{L}{u}{p} > \epsilon$, which implies
  $(\beta H(A) / \epsilon) \regret{L}{u}{p} > \beta$.
  Finally, as $\regret{\ell}{\psi(u)}{p} \leq \beta$, the result follows.
\end{proof}

\subsection{Part 2: separation WLOG}

@@

  Next, observe that, as $L$ is polyhedral, the expected loss $\inprod{p}{L(u)}$ is $\beta$-Lipschitz in $\|\cdot\|_\infty$ for some $\beta>0$.
  \raft{Could write $\beta$ explicitly; it's in my notes}
  Thus, for all $j\in J$, we have
  \begin{align*}
    d_\infty(u_i,\Gamma(p)) \leq \epsilon_j
    &\implies \exists u^*\in\Gamma(p) \|u_j-u^*\|_\infty \leq \epsilon_j
    \\
    &\implies \left| \inprod{p}{L(u_j)} - \inprod{p}{L(u^*)} \right| \leq \beta\epsilon_j
    \\
    &\implies \left| \inprod{p}{L(u_j)} - \risk{L}(p) \right| \leq \beta\epsilon_j~.
  \end{align*}
  Finally, for this $p$, we have
  \begin{align*}
    \inf_{u:\psi(u)\notin\gamma(p)} \inprod{p}{L(u)}
    \leq
    \inf_{j\in J} \inprod{p}{L(u_j)}
    =
    \risk{L}(p)~,
  \end{align*}
  contradicting the calibration of $\psi$.
\end{proof}

@@

  Let $p \in \simplex$ and $u\in\reals^d$ be arbitrary.
  We must have $p\in\hat\gamma_a$ for some $a\in\A$.
  \raft{Using non-emptiness of $\gamma$ here, not that it's a restriction at all}
  By the definition of $\Q_a$, we may write $p = \sum_{q\in\Q_a} \beta_q \, q$ for some $\beta_q \geq 0$, $\sum_{q\in\Q_a} \beta_a = 1$.
  Recall that $\Q_a \subset \hat\gamma_a = \Gamma_{\varphi(a)}$.
  Thus,
  \begin{align*}
    \regret{\ell}{\psi(u)}{p}
    & = \regret{\ell}{\psi(u)}{\textstyle\sum_{q\in\Q_a} \beta_q \, q} &
    \\
    & = \sum_{q\in\Q_a} \beta_q \regret{\ell}{\psi(u)}{q} & \text{Lemma~\ref{lem:regret-linear}}% with $u^*=\varphi(a)$}
    \\
    & = \sum_{q\in\Q_a} \beta_q \alpha_q \regret{L}{u}{q} & \text{Lemma~\ref{lem:regret-bound-fixed-p}}
    \\
    & \leq \left(\max_{q\in\Q_a} \alpha_q\right) \sum_{q\in\Q_a} \beta_q \regret{L}{u}{q} & \text{$\Q_a$ is finite}
    \\
    & \leq \left(\max_{q\in\Q_a} \alpha_q\right) \regret{L}{u}{q} & \text{Lemma~\ref{lem:regret-linear}}% with $u^*=\varphi(a)$}
    \\
    & \leq \alpha \regret{L}{u}{q}~,
  \end{align*}
  completing the proof.
\end{proof}

>>>>

\bo{I don't think existence of an $\epsilon$-separated link $\psi$, for $\epsilon > 0$, implies indirect elicitation, since $(L,\psi)$ could just be totally wrong. Do we want to add into the definition that $(L,\psi)$ indirectly elicits $\ell$?}

@@

\begin{theorem}[Hoffman constant]
  \label{thm:hoffman}
  Given a matrix $A\in\reals^{m\times n}$, there exists some $H(A)\geq 0$, called the \emph{Hoffman constant} (with respect to $\|\cdot\|_\infty$), such that for all $b\in\reals^m$ and all $x\in\reals^n$,
  \begin{equation}
    \label{eq:hoffman}
    d_\infty(x,S(A,b)) \leq H(A) \|(A x - b)_+\|_\infty~,
  \end{equation}
  where $S(A,b) = \{x\in\reals^n \mid A x \leq b\}$ and $(u)_+ \defeq \max(u,0)$ component-wise.
\end{theorem}
\bo{I guess this holds for infinitely many values of $H(A)$. Does ``the'' Hoffman constant usually refer to the inf of that set? If so, we should change our wording slightly to clarify.}

@@

\begin{lemma}[Regret bound for fixed $p$]
  \label{lem:regret-bound-fixed-p}
  Let surrogate $L:\reals^d \to \reals^\Y_+$ and discrete loss $\ell:\R\to\reals^\Y_+$ be given.
  Suppose $(L,\psi)$ indirectly elicits $\ell$, where $\psi:\reals^d\to\R$ is an $\epsilon$-separated link with respect to $\Gamma \defeq \prop{L}$ and $\gamma \defeq \prop{\ell}$ for some $\epsilon>0$.
%  Let $\psi:\reals^d\to\R$ be an $\epsilon$-separated link with respect to $\Gamma \defeq \prop{L}$ and $\gamma \defeq \prop{\ell}$ for some $\epsilon>0$.
  For fixed $p\in\simplex$, there exists some $\alpha\geq 0$ such that
  \begin{equation}
    \label{eq:surrogate-regret-bound-fixed}
    \regret{\ell}{\psi(u)}{p}
    \leq
    \alpha \cdot \regret{L}{u}{p}~,
    % \inprod{p}{\ell(\psi(u))} - \risk{\ell}(p)
    % \leq
    % \alpha \cdot \left( \inprod{p}{L(u)} - \risk{L}(p) \right)~,
  \end{equation}
  for all $u\in\reals^d$.
\end{lemma}

\begin{proof}
  Since $L$ is polyhedral, there exist $a_1,\ldots,a_m \in \reals^d$ and $c\in\reals^m$ such that we may write $\inprod{p}{L(u)} = \max_{1\leq j\leq m} a_j \cdot u + c_j$.
  Let $A \in \reals^{m\times d}$ be the matrix with rows $a_j$, and let $b = \risk{L}(p)\ones - c$, where $\ones\in\reals^m$ is the all-ones vector.
  Then we have
  \begin{align*}
    S(A,b)
    &\defeq \{u\in\reals^d \mid A u \leq b\}
    \\
    &= \{u\in\reals^d \mid A u + c \leq \risk{L}(p)\ones\}
    \\
    &= \{u\in\reals^d \mid \forall i\, (A u + c)_i \leq \risk{L}(p)\}
    \\
    &= \{u\in\reals^d \mid \max_i \;(A u + c)_i \leq \risk{L}(p)\}
    \\
    &= \{u\in\reals^d \mid \inprod{p}{L(u)} \leq \risk{L}(p)\}
    \\
    & = \Gamma(p)~.
  \end{align*}
  Similarly, we have $\max_i\; (A u - b)_i = \inprod{p}{L(u)} - \risk{L}(p) = \regret{L}{u}{p} \geq 0$.
  Thus,
  \begin{align*}
    \|(Au - b)_+\|_\infty
    &= \max_i\; ((Au - b)_+)_i
    \\
    &= \max((Au - b)_1,\ldots,(Au - b)_m, 0)
    \\
    &= \max(\max_i\; (Au - b)_i, \, 0)
    \\
    &= \max_i\; (Au - b)_i
    % \\
    % &= \inprod{p}{L(u)} - \risk{L}(p)
    \\
    &= \regret{L}{u}{p}~.
  \end{align*}

  Now applying Theorem~\ref{thm:hoffman}, we have
  \begin{align*}
    d_\infty(u,\Gamma(p))
    &=    d_\infty(u,S(A,b))
    \\
    &\leq H(A) \|(Au-b)_+\|_\infty
    \\
    &= H(A) \regret{L}{u}{p}~.
  \end{align*}
  Let $\beta \defeq \max_{r\in\R} \regret{\ell}{r}{p}\geq 0$, which is well-defined as $\R$ is a finite set.
  To complete the proof, define $\alpha = \beta H(A)/\epsilon$.
  Given $u\in\reals^d$, if $\regret{\ell}{\psi(u)}{p} = 0$, we are done.
  Otherwise, we have from $\epsilon$-separation that $d_\infty(u,\Gamma(p)) > \epsilon$.
  The above then gives $H(A) \regret{L}{u}{p} > \epsilon$, which implies
  $(\beta H(A) / \epsilon) \regret{L}{u}{p} > \beta$.
  Finally, as $\regret{\ell}{\psi(u)}{p} \leq \beta$, the result follows.
\end{proof}
\bo{Very nice.}
\subsection{Part 2: separation WLOG}

@@

  Next, observe that, as $L$ is polyhedral, the expected loss $\inprod{p}{L(u)}$ is $\beta$-Lipschitz in $\|\cdot\|_\infty$ for some $\beta>0$.
  \raft{Could write $\beta$ explicitly; it's in my notes}
  Thus, for all $j\in J$, we have
  \begin{align*}
    d_\infty(u_i,\Gamma(p)) \leq \epsilon_j
    &\implies \exists u^*\in\Gamma(p) \|u_j-u^*\|_\infty \leq \epsilon_j
    \\
    &\implies \left| \inprod{p}{L(u_j)} - \inprod{p}{L(u^*)} \right| \leq \beta\epsilon_j
    \\
    &\implies \left| \inprod{p}{L(u_j)} - \risk{L}(p) \right| \leq \beta\epsilon_j~.
  \end{align*}
  Finally, for this $p$, we have
  \begin{align*}
    \inf_{u:\psi(u)\notin\gamma(p)} \inprod{p}{L(u)}
    \leq
    \inf_{j\in J} \inprod{p}{L(u_j)}
    =
    \risk{L}(p)~,
  \end{align*}
  contradicting the calibration of $\psi$.
\end{proof}
\bo{Nice. I do think we need to cite that $\Gamma$ takes finitely many values. Also we need to write the definiton of calibration in somewhere.}

@@

  Let $p \in \simplex$ and $u\in\reals^d$ be arbitrary.
  We must have $p\in\hat\gamma_a$ for some $a\in\A$.
  \raft{Using non-emptiness of $\gamma$ here, not that it's a restriction at all}
  By the definition of $\Q_a$, we may write $p = \sum_{q\in\Q_a} \beta_q \, q$ for some $\beta_q \geq 0$, $\sum_{q\in\Q_a} \beta_a = 1$.
  Recall that $\Q_a \subset \hat\gamma_a = \Gamma_{\varphi(a)}$.
  Thus,
  \begin{align*}
    \regret{\ell}{\psi(u)}{p}
    & = \regret{\ell}{\psi(u)}{\textstyle\sum_{q\in\Q_a} \beta_q \, q} &
    \\
    & = \sum_{q\in\Q_a} \beta_q \regret{\ell}{\psi(u)}{q} & \text{Lemma~\ref{lem:regret-linear}}% with $u^*=\varphi(a)$}
    \\
    & = \sum_{q\in\Q_a} \beta_q \alpha_q \regret{L}{u}{q} & \text{Lemma~\ref{lem:regret-bound-fixed-p}}
    \\
    & \leq \left(\max_{q\in\Q_a} \alpha_q\right) \sum_{q\in\Q_a} \beta_q \regret{L}{u}{q} & \text{$\Q_a$ is finite}
    \\
    & \leq \left(\max_{q\in\Q_a} \alpha_q\right) \regret{L}{u}{q} & \text{Lemma~\ref{lem:regret-linear}}% with $u^*=\varphi(a)$}
    \\
    & \leq \alpha \regret{L}{u}{q}~,
  \end{align*}
  completing the proof.
\end{proof}
\bo{Excellent.}
